---
title: "Walkers Report (June 2–6, 2025)"
author: "Jed Ferreras"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(lubridate)
library(ggplot2)
library(dplyr)
library(hms)
library(tidyr)
knitr::opts_chunk$set(echo = TRUE)
```

### Load & Prepare Data
```{r}
data <- read_csv("C:/Users/jed_ferreras/PROMFI/50006-MIMMS.00/Walkers_06022025-06062025.csv")

names(data)[names(data) == "walkers/10 min"] <- "walkers_10_min"
names(data)[names(data) == "focus/work done"] <- "focus_score"

data$clock_time <- as_hms(data$time)
data$date <- as.Date(data$date)
data$rate_per_minute <- data$walkers_10_min / 10
data$rate_per_hour <- data$walkers_10_min * 6
data$hour <- hour(data$clock_time)
data$weekday <- factor(weekdays(data$date), levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
data$period <- ifelse(data$clock_time < as_hms("12:00:00"), "Morning",
                      ifelse(data$clock_time < as_hms("14:00:00"), "Lunch", "Afternoon"))
```

### Daily Summary Stats
```{r}
daily_summary <- data %>%
  group_by(date) %>%
  summarise(
    avg_min = mean(rate_per_minute),
    avg_hour = mean(rate_per_hour),
    max_rate = max(rate_per_minute),
    total_walkers = sum(walkers_10_min),
    avg_focus = mean(focus_score, na.rm = TRUE)
  )

knitr::kable(daily_summary, caption = "Daily Averages and Focus Scores")
```

### Before and After Lunch Periods
```{r}
period_summary <- data %>%
  group_by(date, period) %>%
  summarise(
    avg_rate_min = mean(rate_per_minute),
    avg_focus = mean(focus_score, na.rm = TRUE),
    .groups = "drop"
  )

knitr::kable(period_summary, caption = "Average Rate and Focus by Period")
```

### Peak Time Blocks
```{r}
peak_blocks <- data %>%
  group_by(date) %>%
  top_n(1, rate_per_minute) %>%
  select(date, time, rate_per_minute)

knitr::kable(peak_blocks, caption = "Highest Peak Time Per Day")
```

### Day with Most Foot Traffic
```{r}
busiest_day <- daily_summary %>%
  filter(avg_min == max(avg_min))

knitr::kable(busiest_day, caption = "Day with Highest Average Walkers per Minute")
```

### Extrapolated Daily Walkers and Peak Hour Rate
# Compute max rate per minute and extrapolated full-day total per date
```{r}
extrapolated_summary <- data %>%
  group_by(date) %>%
  summarise(
    observed_minutes = n() * 10,
    total_walkers = sum(walkers_10_min),
    avg_rate_minute = mean(rate_per_minute),
    extrapolated_8hr_total = avg_rate_minute * 60 * 8,
    peak_minute_rate = max(rate_per_minute),
    peak_hourly_rate = round(max(rate_per_minute) * 60, 1),
    .groups = "drop"
  )

# Display table
knitr::kable(
  extrapolated_summary,
  digits = 2,
  caption = "Extrapolated Total Walkers and Peak Hourly Rate (Based on Observed Intervals)"
)
```

### Predicted Walkers by Weekday and Hour
```{r}
pattern_summary <- data %>%
  filter(rate_per_minute > 0) %>%     # Exclude zero traffic times
  group_by(hour) %>%                  # Don't group by weekday anymore
  summarise(
    avg_rate = mean(rate_per_minute),
    .groups = "drop"
  )

knitr::kable(pattern_summary, caption = "Average Walkers per Minute by Hour (All Days, Excluding Zero-Traffic Times)")
```

### Lowest Non-Zero Average Walking Rate
```{r}
min_nonzero_rate <- data %>%
  filter(rate_per_minute > 0) %>%
  summarise(min_rate = min(rate_per_minute)) %>%
  pull(min_rate)

cat("🔻 The lowest non-zero observed walking rate is:", round(min_nonzero_rate, 2), "people per minute\n")
cat("⏱️  That means someone walked by about every", round(60 / min_nonzero_rate), "seconds (at slowest observed rate).\n")
```

### Walkers per Minute Over Each Day
```{r}
ggplot(data, aes(x = clock_time, y = rate_per_minute, color = as.factor(date))) +
  geom_line() +
  labs(title = "Walkers per Minute Over Time", x = "Time", y = "People per Minute", color = "Date") +
  theme_minimal()
```

### Daily Walkers vs Focus Score
```{r}
ggplot(daily_summary, aes(x = avg_min, y = avg_focus)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Daily Walkers vs Focus", x = "Average Walkers per Minute", y = "Focus Score") +
  theme_minimal()
```

### Average Walkers per Minute by Hour
```{r}
ggplot(pattern_summary, aes(x = hour, y = avg_rate)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Average Walkers per Minute by Hour (Excluding Zeros)",
    x = "Hour of Day",
    y = "Avg Walkers per Minute"
  ) +
  theme_minimal()
```

### Summary
** On a daily basis, someone walks by my desk:  **
** • At the **highest rate**: 3 times per minute (once every 20 seconds)  **
** • At the **slowest rate**: once every 2 minutes **
• **Typical interruption studies** report ~1 interruption every 5 minutes **
→ **Conclusion**: Compared to most offices, this environment is *much busier* **